---
title: "Recoding/Labeling Variables"
author: T.E.G.
date: '2018-03-24T00:12:00'
slug: rec
categories: []
tags: []
bibliography: references.bib
---

This is a simple topic. And there are various ways of recoding and (re)labeling variables in `R`. But `rec` function from `sjmisc` package [@sjmisc] works best for me. Give it a try:

```{r rec1, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(strengejacke)

mtcars %>% 
  select(disp, mpg, cyl) %>% 
  rec(cyl,rec= "4=1 [low]; 
                6=2 [mid]; 
                8=3 [high]") %>% 
  to_label(cyl_r)
```

There are other useful functions to transform and recode variables. For example:

```{r rec2, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(strengejacke)

mtcars %>% 
  select(disp, mpg, cyl) %>% 
  rec(cyl,rec= "4=1 [low]; 
                6=2 [mid]; 
                8=3 [high]") %>% 
  split_var(disp, n=4) %>% #splitting variable into n equal sized groups
  dicho(mpg, dich.by = "mean") # dichotomize variable based on a criterion 
```

I recommend checking other strengejacke packages. Especially `sjPlot` which provides a collection of table and plotting functions. Here is the link to the package website: [sjPlot](http://www.strengejacke.de/sjPlot/).

#### **References**

